---
title: "HI Existing Entries Rank 3 Updates"
author: "Madeleine Zoeller"
date: "`r Sys.Date()`"
output: html_document
---

Purpose: Make word clouds from PubMed abstracts to facilitate encyclopedia entry research for the American Botanical Council
#Set up
##Restart
```{r}
rm(list= ls())
```

##Library
```{r}
library(tidyverse)
library(tm)
library(PubMedWordcloud)
library(easyPubMed)
library(readxl)

sessionInfo()
getwd()
```

##Documentation
###PubMedWordcloud info

https://cran.r-project.org/web/packages/PubMedWordcloud/PubMedWordcloud.pdf

###easyPubMed info

https://cran.r-project.org/web/packages/easyPubMed/vignettes/getting_started_with_easyPubMed.html

https://www.data-pulse.com/projects/Rlibs/vignettes/easyPubMed_02_advanced_tutorial.html

Full documentation

https://cran.r-project.org/web/packages/easyPubMed/easyPubMed.pdf


#Existing entries, binned by quality/re-write priority



Existing entries:

-Original length was 108 entries, but I had to separate Bitter Orange, Neroli, Petitgrain and Tea (Green and Black) for coding purposes

-1(highest quality/low rewrite priority) to 3(lowest quality/high rewrite priority)

###Read-in
```{r}
hiranked <- read_excel("Hirankedv2.xlsx")

#will be transformed to lowercase for coding/processing purposes

View(hiranked)

#comName= common name

```

##Ranks
###Rank 1, n=38
```{r}
hiR1 <- hiranked %>% filter(rank== 1)

hiR1
```

###Rank 2, n=32
```{r}
hiR2 <- hiranked %>% filter(rank== 2)

hiR2
```

###Rank 3, n=35
```{r}
hiR3 <- hiranked %>% filter(rank== 3)

hiR3
```

#Literature reviews for Rank 3 items
###Tidy up before starting
```{r}
rm(hiR1)
rm(hiR2)
rm(hiranked)

hiR3 <- hiR3$comName
hiR3 <- tolower(hiR3)
```

#HI Ingredients in Rank 3
##Astaxanthin 
###P1, astaxanthin AND review
```{r}
x = getPMIDsByKeyWords(keys = "astaxanthin AND review", dFrom=2005, dTo=2023)

abstracts = getAbstracts(x)

cleanAbs = cleanAbstracts(abstracts)

pt1_astRev <- plotWordCloud(cleanAbs, min.freq = 4, scale = c(3, 0.3), max.words = 100)

#potential search 2 words: cancer, carotenoid, antioxidant, inflammmation, antiinflammatory oxidative, neurodegenerative, algae, microalgae, liver, nafld, lipid, fatty


#choose: cancer, neurodegenerative, inflammation, antioxidant, carotenoid, liver

```

###P2: astaxanthin AND review AND cancer
```{r}
x = getPMIDsByKeyWords(keys = "astaxanthin AND review AND cancer", dFrom=2005, dTo=2023)

abstracts = getAbstracts(x)

cleanAbs = cleanAbstracts(abstracts, yrWords = exw)

pt2_astRev <- plotWordCloud(cleanAbs, min.freq = 4, scale = c(2, 0.3), max.words = 100)

#new search words: lung, lycopene, diabetes, xanthophylls, fucoxanthin, eye, immune, skin, gastric, cognitive



class(exw)
is.vector(exw)


#words to exclude

exw <- c("disease", "diseases", "study", "treatment", "review", "health", "however", "including", "drugs","disorder","disorders", "used", "use", "using", "several", "can", "well", "also", "research", "elsevier", "studies","can", "shown", "one", "due", "will", "evidence", "current", "increased", "associated", "high", "present", "different", "showed", "play", "found", "reported", "rights", "effects", "effect", "recent")



```
###P3, astaxanthin AND review AND neurodegenerative
```{r}
x = getPMIDsByKeyWords(keys = "astaxanthin AND review AND neurodegenerative", dFrom=2005, dTo=2023)

abstracts = getAbstracts(x)

cleanAbs = cleanAbstracts(abstracts, yrWords = exw)

pt3_astRev <- plotWordCloud(cleanAbs, min.freq = 4, scale = c(2, 0.3), max.words = 100)

#note lack of punctuation
#new words of interest: neuroinflammation, alzheimers, parkinsons, neuroprotective, mitochondrial, stress, gutbrain, renal, eye, apoptosis, mitophagy, klotho, chronic
```

###P4, astaxanthin AND review AND neurodegenerative AND apoptosis
```{r}
x = getPMIDsByKeyWords(keys = "astaxanthin AND review AND neurodegenerative AND apoptosis", dFrom=2005, dTo=2023)

abstracts = getAbstracts(x)

cleanAbs = cleanAbstracts(abstracts, yrWords = exw)

pt4_astRev <- plotWordCloud(cleanAbs, min.freq = 4, scale = c(2, 0.3), max.words = 100)

#note lack of punctuation
#new words of interest: neuroinflammation, alzheimers, parkinsons, neuroprotective, mitochondrial, stress, gutbrain, renal, eye, apoptosis, mitophagy, klotho, chronic

#fibrosis, pikakt, fgf

```
###P5,astaxanthin AND review AND future AND directions 2020-present

```{r}
x = getPMIDsByKeyWords(keys = "astaxanthin AND review AND future AND directions", dFrom=2020, dTo=2023)

#x2 = list(x)
#use as list and feed into code example from easyPubMed above?
abstracts = getAbstracts(x)

cleanAbs = cleanAbstracts(abstracts, yrWords = exw)

pt5_astRev <- plotWordCloud(cleanAbs, min.freq = 4, scale = c(2, 0.3), max.words = 100)

#Relevant words: biorefinery, sustainable, solvents, pluvialis
```



```{r}
hiR3
```

###Astaxanthin and Safety
```{r}

x = getPMIDsByKeyWords(keys = "astaxanthin AND safety", dFrom=2020, dTo=2023)

abstracts = getAbstracts(x)

cleanAbs = cleanAbstracts(abstracts, yrWords = exw)

pt6_astSafe <- plotWordCloud(cleanAbs, min.freq = 4, scale = c(3, 0.6), max.words = 100)

#look into
#embryos, glucose, haematococcus, toxicity, zebrafish, knee, bioavailability, concentrations, extraction, shrimp
```


###Astaxanthin and Sustainability
```{r}
x = getPMIDsByKeyWords(keys = "astaxanthin AND safety", dFrom=2020, dTo=2023)

abstracts = getAbstracts(x)

cleanAbs = cleanAbstracts(abstracts, yrWords = exw)

pt7_astSustain <- plotWordCloud(cleanAbs, min.freq = 4, scale = c(3, 0.6), max.words = 100)

#look into
#pluvialis, knee, shrimp, 
```



##Beet
###P1, beet AND review
```{r}

x = getPMIDsByKeyWords(keys = "beet AND review", dFrom=2005, dTo=2023)

abstracts = getAbstracts(x)

cleanAbs = cleanAbstracts(abstracts)

pt1_beetRev <- plotWordCloud(cleanAbs, min.freq = 4, scale = c(1, 0.3), max.words = 100)


```

###P2, beet AND review (minus previous exw list)
```{r}

x = getPMIDsByKeyWords(keys = "beet AND review", dFrom=2005, dTo=2023)

abstracts = getAbstracts(x)

cleanAbs = cleanAbstracts(abstracts, yrWords = exw)

pt2_beetRev <- plotWordCloud(cleanAbs, min.freq = 4, scale = c(1, 0.3), max.words = 100)

#acne vulgaris, betalains, nitrate, blood, cardiovascular, beetroot, sugar, sulopenem, antioxidant, regulatory, virus, acid, pectin

#adding to excluded words list (exw)
#including, related, following, identified, human, patients, important, factors, beneficial, published, benefits, properties, new

#borderline exw....various, many
```
###new set of excluded words
```{r}
exw <- c("disease", "diseases", "study", "treatment", "review", "health", "however", "including", "drugs","disorder","disorders", "used", "use", "using", "several", "can", "well", "also", "research", "elsevier", "studies","can", "shown", "one", "due", "will", "evidence", "current", "increased", "high", "present", "different", "showed", "play", "found", "reported", "rights", "effects", "effect", "recent"
         #start of beet additions
         ,"including","related", "following", "identified", "human", "patients", "important", "factors", "beneficial", "published", "benefits", "properties", "new")
```

###P3, beet AND review AND acne
```{r}
x = getPMIDsByKeyWords(keys = "beet AND review AND acne", dFrom=2005, dTo=2023)

abstracts = getAbstracts(x)

cleanAbs = cleanAbstracts(abstracts, yrWords = exw)

pt3_beetRev <- plotWordCloud(cleanAbs, min.freq = 4, scale = c(3, 0.3), max.words = 100)

#words of interest: skin, isotretinoin, cancer, sebocyte, keratinocyte, immunomodulatory, inflammation, rosacea, antibiotics, psoriasis, curcumin, oral, lesions, adp, cosmetic, peptides, amps, systemic, insulinotropic, combination, milk, sabaceous, retinoids, neutrophils

#decided to remove associated from exw list, this indicates the type of evidence that exists
```

##Bergamot
###excluded words
```{r}
exw <- c("disease", "diseases", "study", "treatment", "review", "health", "however", "including", "drugs","disorder","disorders", "used", "use", "using", "several", "can", "well", "also", "research", "elsevier", "studies","can", "shown", "one", "due", "will", "evidence", "current", "increased", "high", "present", "different", "showed", "play", "found", "reported", "rights", "effects", "effect", "recent"
         #start of beet additions
         ,"including","related", "following", "identified", "human", "patients", "important", "factors", "beneficial", "published", "benefits", "properties", "new"
         #start of bergamot additions
         , "studies")
```

###P1, bergamot AND review
```{r}

x = getPMIDsByKeyWords(keys = "bergamot AND review", dFrom=2005, dTo=2023)

abstracts = getAbstracts(x)

cleanAbs = cleanAbstracts(abstracts)

pt1_bergRev <- plotWordCloud(cleanAbs, min.freq = 4, scale = c(3, 0.3), max.words = 100)

#words of interest: linalool, carvacrol, cholesterol, citrus, volatile, bergamia, beo, cardiovascular, aromatherapy, antiinflammatory, yeast, lipoprotein
```




####For later PubMed bergamot and review csv download
```{r}
bergCSV <- read.xl("")
```



###P2, bergamot AND review, exw
```{r}

x = getPMIDsByKeyWords(keys = "bergamot AND review", dFrom=2005, dTo=2023)

abstracts = getAbstracts(x)

cleanAbs = cleanAbstracts(abstracts, yrWords = exw)

pt1_bergRev <- plotWordCloud(cleanAbs, min.freq = 4, scale = c(3, 0.3), max.words = 100)


#words of interest: bergamia, LDL, stress, naringenin, yeast, juice, safety, statins, lipid lowering, cancer, anticancer, auraantium, antimicrobial, neuropathic, HDLC, carvacrol, linalool, rice, eos, berberine, anxiety, antiinflammatory, artichoke, orange, cholesterol, volatile, beo, red, citrus, metabolism
```
##Trying something else


-maybe I need to separate this into steps?
-starting with vector...or list? of my HI ingredients 
-then I want a ____loop? something in the apply family? that will write out the keys = part "[element in list or vector of HI ingredients] AND review" 

What if I try to pipe these together and then try to use a loop to go through my list?
```{r}
#remember to add in yrWords = exc  for clean abstracts in another version of this 

#going to test this with a much smaller search b/c I am trying this on the train (did not need to it turns out)
x = getPMIDsByKeyWords(keys = "bergamot AND review", dFrom=2005, dTo=2023)

#maybe each of the arguments above for getPMIDsByKeyWords() could have its own column?
#dataframe ...colNames = keys(created before, example value for the Bergamot row, e.g."bergamot AND review"), dFrom, dTo


#----(the x= part?) 

#would need a way to do this that does not just write over each existing version of x?

#also would prefer if it can name each x equivalent in way that shows it is a list of PMIDs for a certain HI ingredient, e.g. beetPMIDs

#maybe I can use mutate?


#below worked!
x %>% getAbstracts() %>% cleanAbstracts() %>% plotWordCloud(min.freq = 4, scale = c(3, 0.3), max.words = 100)

```
Words version of the loop I would want (or maybe lapply, that might be faster?)

-For each value in list or vector of HI ingredient names, feed that value into the keys= argument within the function getPMIDsByKeywords and create a new column with the output of the function getPMIDsByKeywords in the same dataframe as my HI list or vector; then, pass the x equivalent for each value (common name)

dataframe....list?- first column is just the HI common name, the next column is "[HI item] AND review" and the third column is equivalent to x above (the output of getPMIDsByKeywords)

-maybe that list or vector should just say "[HI item] AND review"

##Cassava
###P1, Cassava and review
```{r}
x = getPMIDsByKeyWords(keys = "cassava AND review", dFrom=2005, dTo=2023)

abstracts = getAbstracts(x)

cleanAbs = cleanAbstracts(abstracts)

pt1_cassRev <- plotWordCloud(cleanAbs, min.freq = 4, scale = c(3, 0.3), max.words = 100)

#words of interest: security, genetic, maize, potato, starch, roots, cyanide, manihot, africa, alzheimer's
```
###P2, Cassava and review, exw
```{r}
x = getPMIDsByKeyWords(keys = "cassava AND review", dFrom=2005, dTo=2023)

abstracts = getAbstracts(x)

cleanAbs = cleanAbstracts(abstracts, yrWords = exw)

pt2_cassRev <- plotWordCloud(cleanAbs, min.freq = 4, scale = c(3, 0.3), max.words = 100)

#words of interest: cyanide, alzheimer's, drug, resistance, sustainable, environmental, virus, risk, flour, konzo, wheat, rice, sweet, tropical, tuber, staple

```

###excluded words
```{r}
exw <- c("disease", "diseases", "study", "treatment", "review", "health", "however", "including", "drugs","disorder","disorders", "used", "use", "using", "several", "can", "well", "also", "research", "elsevier", "studies","can", "shown", "one", "due", "will", "evidence", "current", "increased", "high", "present", "different", "showed", "play", "found", "reported", "rights", "effects", "effect", "recent"
         #start of beet additions
         ,"including","related", "following", "identified", "human", "patients", "important", "factors", "beneficial", "published", "benefits", "properties", "new"
         #start of bergamot additions
         , "studies",
         #cassava additions
         "included", "many", "countries", "results", "information", "need", "understanding", "provide", "developing", "people", "knowledge", "reserved", "among","method","authors","key","literature")
```


##Coconut

###P1, Coconut and review
```{r}
x = getPMIDsByKeyWords(keys = "coconut AND review", dFrom=2005, dTo=2023)

abstracts = getAbstracts(x)

cleanAbs = cleanAbstracts(abstracts)

pt1_coconRev <- plotWordCloud(cleanAbs, min.freq = 4, scale = c(3, 0.3), max.words = 100)
```
###P2, Coconut and review, exw
```{r}
x = getPMIDsByKeyWords(keys = "coconut AND review", dFrom=2005, dTo=2023)

abstracts = getAbstracts(x)

cleanAbs = cleanAbstracts(abstracts, yrWords = exw)

pt2_coconRev <- plotWordCloud(cleanAbs, min.freq = 4, scale = c(3, 0.3), max.words = 100)

#words of interest: cardiovascular, cholesterol, antioxidant, environmental, herbs, olive, waste, poria


#try to add ggsave because it isn't allowing me to right click the way I could before....

x %>% getAbstracts() %>% cleanAbstracts() %>% plotWordCloud(min.freq = 4, scale = c(3, 0.3), max.words = 100)

#words of interest: waste, skin, cardiovascular, olive (presumably comparisons with olive oil), poria, rice, antioxidant, preventative, cholesterol, saturated, environmental, energy 
```

###excluded words
```{r}
exw <- c("disease", "diseases", "study", "treatment", "review", "health", "however", "including", "drugs","disorder","disorders", "used", "use", "using", "several", "can", "well", "also", "research", "elsevier", "studies","can", "shown", "one", "due", "will", "evidence", "current", "increased", "high", "present", "different", "showed", "play", "found", "reported", "rights", "effects", "effect", "recent"
         #start of beet additions
         ,"including","related", "following", "identified", "human", "patients", "important", "factors", "beneficial", "published", "benefits", "properties", "new"
         #start of bergamot additions
         , "studies",
         #cassava additions
         "included", "many", "countries", "results", "information", "need", "understanding", "provide", "developing", "people", "knowledge", "reserved", "among","method","authors","key","literature",
         #coconut addition
         "therefore", "methods")
```


##Cranberry
##Cucumber
##Damask rose
##English walnut
##Eucalyptus
##Fennel
##Frankincense
##Grape leaf
##Grapefruit
##Guar gum
##Hops
##Hyssop
##Jojoba
##Madder
##Mallows
##Tangerine
##Marjoram
##Mugwort
##Myrrh
##Patchouli
##Peppermint
##Rice
##Roman chamomile
##Sandalwood
##Sweet almond
##Black tea
##Turmeric
##Vetiver
##Wheat germ
##Ylang ylang